---
title: "Solar Radiation"
format: html
editor: visual
bibliography: references.bib
---

```{r}
library(azmetr)
library(tidyverse)
library(units)
daily <- az_daily(start_date = "2023-01-01")
hourly <- az_hourly(start_date_time = "2023-08-19 00")

daily <- daily |> az_add_units()
hourly <- hourly |> az_add_units()
```

Looking for MJ/m^2^ or possibly W/m^2^ which includes time in the denominator

```{r}
(daily$sol_rad_total[1] / set_units(1, "d") )|> set_units("W/m^2")
```

## Parameter estimates

We can get a bunch of data about solar radiation with `fishmethods::astrocalc4r()`

```{r}
library(fishmethods)

hourly_sol <- 
  hourly |> 
  left_join(station_info) |> 
  add_column(tz = -7) |> 
  reframe(
    astrocalc4r(
      day = mday(date_datetime),
      month = month(date_datetime),
      year = year(date_datetime),
      hour = hour(date_datetime),
      timezone = tz,
      lat = latitude,
      lon = longitude,
      withinput = FALSE,
      seaorland = "continental"
    )
  ) 

hourly_sol <- 
  hourly |> 
  select(meta_station_id,
    meta_station_name,
    date_datetime,
    date_doy,
    sol_rad_total) |> 
  bind_cols(hourly_sol) |> 
  rename(par = PAR) |> 
  mutate(par = set_units(par, "lux"))
```

## PAR (not what we want)

This gives us PAR in units of lux, which unfortunately is too small of a slice of wavelenghts to be comparable to the total solar radiation data.
Here is how it can be converted to comparable units, just in case that is useful later.

```{r}
sunlight <- set_units(0.0079, "(W/m^2) / Lux") # approximate.
# https://physics.stackexchange.com/questions/135618/rm-lux-and-w-m2-relationship

hourly_sol <- 
  hourly_sol |> 
  mutate(par = par * sunlight) |> 
  mutate(par = set_units(par, "MJ m-2 hr-1") * set_units(1, "hr"))
```

Now we can plot this maybe

```{r}
hourly_sol |> 
  ggplot(aes(x = date_datetime)) +
  geom_line(aes(y = sol_rad_total)) +
  geom_line(aes(y = par), color = "red") +
  facet_wrap(~meta_station_id)
```

> `astrocalc4r` calculates PAR for wavelengths between 400-700 nm.
> Calculations for other wavelengths can be carried out by modifying the code to use parameters from Frouin et al. (1989).

So, it might be possible to dig into the code and modify it.

## Extraterrestrial Radiation

We can estimate radiation to the outside of the atmosphere using the zenith and solar constant

```{r}
zenith_to_sol_rad <- function(zenith){
  foo <-  function(zenith) {

    #convert to radians
    zenith_rad <- zenith * pi / 180
    #cos zenith angle
    cosz <- cos(zenith_rad)
    #assume 0 radiation when angle ≥ 90
    if(zenith >= 90) {
      radiation <- 0
    } else {
      #multiply by "solar constant" 
      radiation <-  cosz * 1358.2 #W/m^2 for 250-4000nm range
    }
    radiation
  }
  radiation <- purrr:::map_dbl(zenith, foo)

  set_units(radiation, "W/m^2")
}

hourly_sol <- 
  hourly_sol |> 
  mutate(sol_potential = (zenith_to_sol_rad(zenith) * set_units(1, "hr")) |> 
           set_units("MJ/m^2"))
```

```{r}
ggplot(hourly_sol, aes(x = date_datetime)) +
  geom_line(aes(y = sol_rad_total)) +
  geom_line(aes(y = sol_potential), color = "red") +
  facet_wrap(~meta_station_id)
```

These estimates are obviously too high because they aren't taking into account the atmosphere!

Next steps are either:

1.  Figure out how to mathematically account for atmosphere (e.g., by getting elevation data for sites)
2.  Empirically estimate some correction factor for each site using all the data.
3.  Modify `astrocalc4r` to work for a greater range of wavelengths

## Digging into `astrocalc4r`

```{r}
astrocalc4r
```
